% !TEX TS-program = pdflatex
% !TEX encoding = UTF-8 Unicode
\documentclass[12pt]{article}
\usepackage[utf8]{inputenc} 
\usepackage[a4paper,left=3cm,right=2cm,top=2cm,bottom=2cm]{geometry}
\usepackage[swedish]{babel}
%\usepackage{tgtermes}
\begin{document}
\pagenumbering{gobble}% Remove page numbers (and reset to 1)
\clearpage\thispagestyle{empty}

\section*{ETSA02\\Role assignment and collaboration contract Group \texttt{00}}
\subsection*{Collaboration rules}
We all agree to comply with the following rules to ensure a positive learning environment for everyone involved. We will:
\begin{enumerate}
\item listen to each other's ideas with respect.
\item do our work as best as we can and ask for help if we need it.
\item do our work in time to help the team meet the project deadlines.
\item arrive on time and well prepared for group meetings.
\item contribute to an inclusive and respectful team.
\item ... \emph{feel free to add more}
\end{enumerate}

If someone on our team breaks one or more of our agreements, the team may have a meeting and ask the person to follow our agreement. If the person still breaks an agreement, we will ask our project supervisor to help find a solution.

\subsection*{Agreement and role assignment}
We, the undersigned, all agree to comply with the above rules. Furthermore, we agree to assign roles to the team members as follows:

\vspace{1em}

\begin{tabular}{l p{4cm} l p{4cm}}
\emph{Role} & \emph{Name} & \emph{Stil-ID} & \emph{Signature} \\
\\ Project manager & \dotfill & \dotfill & \dotfill \\
\\ Requirements engineer & \dotfill & \dotfill & \dotfill \\
\\ Development lead & \dotfill & \dotfill & \dotfill \\
\\ Test lead & \dotfill & \dotfill & \dotfill \\
\\ Sales engineer & \dotfill & \dotfill & \dotfill \\
\\ Domain expert & \dotfill & \dotfill & \dotfill \\
\end{tabular}

\vspace{1em }
\subsection*{Approved by project supervisor}

\begin{tabular}{l p{8cm}}
\\ Name and signature &  \dotfill \\
\\ Date & \dotfill \\
\end{tabular}

\end{document}
